---
title: "EDS 230 Assignment 3: Almond Profit Model Sensitivity Analysis"
author: "Elmera Azadpour, Mia Forsline, Alex Vand"
date: '2022-04-19'
output:
  pdf_document: default
  html_document: default
---
# Set up 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE, 
                      warning = FALSE)

#install packages if necessary, then load libraries
if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}

librarian::shelf(
  ggpubr,
  here,
  tidyverse,
  tidyverse)
```

# Introduction 

# 1. Develop a profit model for almond yield 

```{r}
#read in R script to compute almond profit 
#more details about the profit model can be found in the .R file 
source(here("functions", "compute_profit_from_almonds.R"))

#read in other necessary R scripts 

# 1. read in R script to compute almond yield anomaly 
source(here("functions", "almond_yield_anomaly_annual.R"))

# 2. read in R script to compute net present value of almonds 
source(here("functions", "compute_npv.R"))
```

Test the `almond_yield_anomaly_annual()` function
- save the output as a CSV 

```{r}
# read in the data from .txt file
clim_raw <- read.csv(file = here("data", "clim.txt"), 
                     header = T, 
                     sep = "")

# clean data 
clim_data <- clim_raw %>% 
             janitor::clean_names() %>% 
             mutate(d = lubridate::as_date(d), #convert d column to Date format rather than a character 
                    year = lubridate::year(d))

# set coefficient parameter values 
temp_coeff1 <- -0.015
temp_coeff2 <- -0.0046
precip_coeff1 <- -0.07
precip_coeff2 <- 0.0043
constant <- 0.28

# calculate annual almond yield for each water year (wy)
annual_almond_yield <- cbind(unique(clim_data$wy), almond_yield_anomaly_annual(clim_data)) %>%  #create a matrix/array of dates and annual almond yields
  as.data.frame() %>% #transform into a data frame 
  rename(year = V1,
         yield = V2)

#save output as a CSV in the data folder 
write_csv(x = annual_almond_yield, path = here("data", "annual_almond_yield.csv"))
```


Test the `compute_npv()` function 

```{r}
npv_test <- round(compute_npv(value = 10, time = 2, discount = 0.12), digits = 2)
```

With a discount rate of 12% over 2 years, a current value of $10 will have a net present value (NPV) of approximately $`r npv_test`. 

# 2. Do a simple informal sensitivity analysis of almond yield profit using at least 2 parameters 

- Similar to the in-class example, we plan to conduct a sensitivity analysis assuming +/- 15% uncertainty in the current price of almonds (`price`) and discount rate (`discount`) 
- We assume a default almond price of $3,000/ton 
- We assume a default discount rate of 0.12 
- We begin by sampling 300 times from a uniform distribution 

```{r}
#parameter defaults 
almond_price_default <- 3000 #$1.47/lb * 2000lb = approximately $3000/ton 
discount_rate_default <- 0.12 

#deviation %
deviation = 0.15 

#number of samples
nsamples = 300

#sample a uniform distribution from the price default
price <- runif(min = almond_price_default - (deviation * almond_price_default), 
               max = almond_price_default + (deviation * almond_price_default),
               n=nsamples)

discount = rnorm(mean=0.6, sd = 0.1, n=nsamples)

#bind price_thresh and discount_rate into a dataframe 
parameters <- cbind.data.frame(discount, price)
```


```{r}
#read in CSV of output from annual almond yield anomaly function 
annual_almond_yield <- read_csv(here("data", "annual_almond_yield.csv"))

#note that parameters column names must match the input parameter names in the compute_profit_from_almonds function
## testing creation of results by using static values for almond_yield_anomaly and year 
results <- parameters %>% 
  pmap(compute_profit_from_almonds,
       almond_yield_anomaly = annual_almond_yield$yield,
       year = annual_almond_yield$year)

#check the results 
results[[1]]
#length(results) - should be a length of 300 
```


```{r}
# now we can extract data from the results list created above 
mean <- map_df(results, #list we want to extract data from 
               `[`, # `[` is a function that means extract 
               c("year", 
                 "net", 
                 "netpre")) #select the columns from results we want to extract 

# and we can add the parameter values for each run so we know what parameters gave us which mean value
mean <- cbind.data.frame(mean, parameters)
```

# 3. Create a single graph of the results 

- remember, for our sensitivity analysis, we chose to vary the price and discount rate to determine how almond yield changed over time 

```{r, results = FALSE}
# plot - pick one of the 2 parameter as a color

p1 <- ggplot(data = mean, aes(x = netpre, y = price, color = discount)) + 
  geom_point(cex=2, alpha = 0.2) + # alpha = opacity
  labs(y = "Price", 
       x = "NPV",
       color = "Discount Rate")
  theme_classic()

p2 <- ggplot(data = mean, aes(x = discount, y = price, color = netpre)) + 
  geom_point(cex=2, alpha = 0.1) + # alpha = opacity
  labs(y = "Price", 
       x = "Discount Rate",
       color = "NPV") + 
  theme_classic()

ggarrange(p1,p2)
```





```{r, results = FALSE}
# plot - the 2 parameters as independent variables on x-axis
# NOTE THAT THESE PLOTS LOOK THE SAME AS THE ONES ABOVE - does this make sense?

p3 <- ggplot(data = mean, aes(x = netpre, y = price, color = discount)) + 
  geom_point(cex=2, alpha = 0.2) + # alpha = opacity
  labs(y = "Price", 
       x = "NPV", # parameter 1
       color = "Discount")
  theme_classic()

p4 <- ggplot(data = mean, aes(x = discount, y = price, color = netpre)) + 
  geom_point(cex=2, alpha = 0.2) + # alpha = opacity
  labs(y = "Price", 
       x = "Discount", # parameter 2
       color = "NPV") + 
  theme_classic()

ggarrange(p3,p4)

ggarrange(p1,p2,p3,p4) #to compare all 4 graphs
```




# 4. Output the graph as a stand alone image

- save as .png

```{r}
#save image in assignments directory
ggsave("assignment3_graph.png",
       plot = ggarrange(p1,p2))
```


